\vspace{-1em}Guidance, please! Towards a framework for RDF-based constraint languages.

### 2015-07-07 paper/paper 2/dc-2015.tex
section 2.2

A \emph{constraint language} is a language which is used to formulate constraints.
The \emph{W3C} working group defines \emph{RDF constraint} as a component of a schema what needs to be satisfied\footnote{\url{https://www.w3.org/2014/data-shapes/wiki/Glossary}}.
As there are already five promising constraint languages, 
our purpose was not to invent a new language.
As there is no high-level declarative language which meets all RDF validation requirements,
we rather developed a very simple lightweight vocabulary,
consisting of only three classes, three object properties, and three data properties, 
which is universal enough to describe constraints of any constraint type.
We call this vocabulary the \emph{RDF Constraints Vocabulary (RDF-CV)}\footnote{Online available at: \url{https://github.com/boschthomas/RDF-Constraints-Vocabulary}} whose conceptual model is shown in Figure \ref{fig:RDF-CV-conceptual-model}.
\begin{figure}[H]
	\centering
		\includegraphics[width=0.60\textwidth]{images/RDF-CV-conceptual-model.png}
	\caption{\emph{RDF Constraints Vocabulary (RDF-CV)} Conceptual Model}
	\label{fig:RDF-CV-conceptual-model}
\end{figure}

We classify sets of constraints according to four dimensions.
For each constraint type, we evaluated \cite{BoschNolleAcarEckert2015} in which set of constraints their instances are contained.
%In the following, we state the rounded number of associated constraint types in brackets.
%\begin{itemize}
  %\item \emph{Universality:} \emph{specific constraints} vs. \emph{generic constraints}
	%\item \emph{Complexity:} \emph{simple constraints} vs. \emph{complex constraints}
	%\item \emph{Context:} \emph{property constraints} vs. \emph{class constraints}
	%\item \emph{DL Expressivity:} \emph{constraints expressible in DL} vs. \emph{constraints not expressible in DL}
%\end{itemize}
%(1) \emph{Universality:} \emph{specific constraints} vs. \emph{generic constraints},
%(2) \emph{Complexity:} \emph{simple constraints} vs. \emph{complex constraints},
%(3) \emph{Context:} \emph{property constraints} vs. \emph{class constraints}, and
%(4) \emph{DL Expressivity:} \emph{constraints expressible in DL} vs. \emph{constraints not expressible in DL}.
\textbf{\emph{Universality}:}
Constraints are expressible either generically (\emph{generic constraints}) using the \emph{RDF-CV}   
or specifically (\emph{specific constraints}) by domain-specific constraint languages. 
As we express constraints of each constraint type in form of \emph{generic constraints} conforming to the \emph{RDF-CV} \cite{BoschNolleAcarEckert2015},
we show that constraints of any constraint type can be described generically.
\textbf{\emph{DL Expressivity}:}
The \emph{RDF-CV} allows to describe 
constraints which are \emph{expressible in DL} (constraints of this set are instantiated from 64\% of the overall 81 constraint types) and which are \emph{not expressible in DL} (36\%). 
%In case a constraint type is \emph{expressible in DL}, we determined which \emph{DL} constructs are needed to express the constraint type.
\textbf{\emph{Complexity}:}
\emph{Simple constraints} (60\% of the constraint types) denotes the set of atomic constraints. 
%which can be instantiated from almost 2/3 of the constraint types.
\emph{Complex constraints} (26\%) is the set of constraints which are created out of \emph{simple} and/or other \emph{complex constraints}.
%and which are instances of 1/4 of the constraint types.
%In case \emph{complex constraints} are expressible in \emph{DL}, 
%\emph{DL} statements representing \emph{complex constraints} contain \emph{DL} statements standing for atomic and/or composed constraints. 
Constraints of almost 14\% of the constraint types are \emph{complex constraints}
which can be simplified and therefore formulated as \emph{simple constraints} when using them in terms of syntactic sugar.
\textbf{\emph{Context}:}
\emph{Simple constraints} are applied on properties (\emph{property constraints}, 60\%),
on classes (\emph{class constraints}, 25\%), or
on properties and classes (\emph{property and class constraints}, 15\%).
%2/3 of all constraint types are types of \emph{property constraints}, 
%1/5 of \emph{class constraints}, and 
%approx. 10\% of \emph{property and class constraints}.
\emph{Simple constraints} must hold for individuals of their \emph{context classes}.
As \emph{context classes} of \emph{simple constraints} are reused within \emph{complex constraints},
it does not make sense to have terms standing for \emph{simple} and \emph{complex constraints} within the \emph{RDF-CV}. 
As a consequence, the distinction of \emph{property} and \emph{class constraints} is sufficient to generically describe constraints of all constraint types.
%\tb{use RDF-CV / not: expressible in RDF-CV / RDF-CV is a vocabulary, not a language}
Altogether, instances of the most of the constraint types are \emph{simple constraints on properties} which are \emph{expressible in DL}. 
%Thus, the majority of constraint types are directly and relatively easy formulated in form of \emph{simple constraints} which can be mapped to equivalent \emph{DL} constructs.

>>>>

A \emph{constraint language} is a language which is used to formulate constraints.
The W3C working group defines \emph{RDF constraint} as a component of a schema what needs to be satisfied\footnote{\url{https://www.w3.org/2014/data-shapes/wiki/Glossary}}.
As there are already five promising constraint languages, 
our purpose is not to invent a new language.
As there is no high-level declarative language which meets all RDF validation requirements,
we rather developed a very simple lightweight vocabulary,
consisting of only three classes, three object properties, and three data properties, 
which is universal enough to describe constraints of any constraint type.
We call this vocabulary the \emph{RDF Constraints Vocabulary (RDF-CV)}\footnote{Online available at: \url{https://github.com/boschthomas/RDF-Constraints-Vocabulary}} whose conceptual model is shown in Figure \ref{fig:RDF-CV-conceptual-model}.
\begin{figure}[H]
	\centering
		\includegraphics[width=0.60\textwidth]{images/RDF-CV-conceptual-model.png}
	\caption{\emph{RDF Constraints Vocabulary (RDF-CV)} Conceptual Model}
	\label{fig:RDF-CV-conceptual-model}
\end{figure}

We classify sets of constraints according to four dimensions.
For each constraint type, we evaluated \cite{BoschNolleAcarEckert2015} in which set of constraints their instances are contained.
%In the following, we state the rounded number of associated constraint types in brackets.
%\begin{itemize}
  %\item \emph{Universality:} \emph{specific constraints} vs. \emph{generic constraints}
	%\item \emph{Complexity:} \emph{simple constraints} vs. \emph{complex constraints}
	%\item \emph{Context:} \emph{property constraints} vs. \emph{class constraints}
	%\item \emph{DL Expressivity:} \emph{constraints expressible in DL} vs. \emph{constraints not expressible in DL}
%\end{itemize}
%(1) \emph{Universality:} \emph{specific constraints} vs. \emph{generic constraints},
%(2) \emph{Complexity:} \emph{simple constraints} vs. \emph{complex constraints},
%(3) \emph{Context:} \emph{property constraints} vs. \emph{class constraints}, and
%(4) \emph{DL Expressivity:} \emph{constraints expressible in DL} vs. \emph{constraints not expressible in DL}.

\textbf{\emph{Universality}:}
Constraints are expressible either generically (\emph{generic constraints}) using the \emph{RDF-CV}   
or specifically (\emph{specific constraints}) by domain-specific constraint languages. 
As we express constraints of each constraint type in form of \emph{generic constraints} conforming to the \emph{RDF-CV} \cite{BoschNolleAcarEckert2015},
we show that constraints of any constraint type can be described generically.

\textbf{\emph{DL Expressivity}:}
The \emph{RDF-CV} allows to describe 
constraints which are either \emph{expressible in DL} (64\%) or which are \emph{not expressible in DL} (36\%). Constraints which are expressible in DL are instantiated from 64\% of the overall 81 constraint types. On the other hand, constraints which are not expressible in DL are assigned to 36\% of all constraint types.
%In case a constraint type is \emph{expressible in DL}, we determined which \emph{DL} constructs are needed to express the constraint type.

\textbf{\emph{Complexity}:}
\emph{Simple constraints} (60\% of the constraint types) denotes the set of atomic constraints. 
%which can be instantiated from almost 2/3 of the constraint types.
\emph{Complex constraints} (26\%) is the set of constraints which are created out of \emph{simple} and/or other \emph{complex constraints}.
%and which are instances of 1/4 of the constraint types.
%In case \emph{complex constraints} are expressible in \emph{DL}, 
%\emph{DL} statements representing \emph{complex constraints} contain \emph{DL} statements standing for atomic and/or composed constraints. 
Constraints of almost 14\% of the constraint types are \emph{complex constraints}
which can be simplified and therefore formulated as \emph{simple constraints} when using them in terms of syntactic sugar.

\textbf{\emph{Context}:}
\emph{Simple constraints} are applied on properties (\emph{property constraints}, 60\%),
on classes (\emph{class constraints}, 25\%), or
on properties and classes (\emph{property and class constraints}, 15\%).
%2/3 of all constraint types are types of \emph{property constraints}, 
%1/5 of \emph{class constraints}, and 
%approx. 10\% of \emph{property and class constraints}.
\emph{Simple constraints} must hold for individuals of their associated \emph{context classes}.
As \emph{context classes} of \emph{simple constraints} are reused within \emph{complex constraints},
it does not make sense to have terms standing for \emph{simple} and \emph{complex constraints} within the \emph{RDF-CV}. 
As a consequence, the distinction of \emph{property} and \emph{class constraints} is sufficient to generically describe constraints of all constraint types.
%\tb{use RDF-CV / not: expressible in RDF-CV / RDF-CV is a vocabulary, not a language}

Altogether, instances of the most of the constraint types are \emph{simple constraints on properties} which are \emph{expressible in DL}. 
%Thus, the majority of constraint types are directly and relatively easy formulated in form of \emph{simple constraints} which can be mapped to equivalent \emph{DL} constructs.
